Ribosome accumulation during early phase resistance training

### 2020-08-24 manuscript.Rmd
update manuscript

Our skeletal muscles not only constitutes our locomotive apparatus but also serve as our primary storage site of amino acids. This makes the musculoskeletal system a critical target for interventions that promotes health across the lifespan [@RN763]. The beneficial effects of resistance-training (RT) are in large part mediated through exercise-induced changes in the balance between muscular protein breakdown and synthesis, with successful RT leading to net protein accretion and muscle hypertrophy. A traditional view of exercise-induced muscle hypertrophy has been that muscle protein accumulates as a consequence of repeated episodes of elevated post-exercise protein synthesis in the acute phase, resulting in a net-positive protein balance over time. Indeed, one bout of resistance exercise acutely up-regulates skeletal muscle protein synthesis up to 48-hr after exercise [@RN786] and subsequently, repeated bouts would lead to accumulation of muscle protein over time [@RN1272; @RN1642]. 
In recent years, this view has been supplemented with evidence showing that chronic RT also leads to an increased resting synthetic rate of muscle protein [@RN1521; @RN1866; @RN1897]. An increase in the resting protein synthetic rate has been postulated to be related to an increased translational capacity, i.e. accumulation of ribosomes [@RN1897; @RN1912], a notion supported by observations relating exercise-induced changes in total RNA, a proxy marker for ribosome abundance, to muscle hypertrophy [@RN1644, @RN1755]. Furthermore, in vitro studies shows that direct inhibition of ribosomal RNA (rRNA) transcription or inhibition of up-stream transcription factors leading to halted biosynthesis of ribosomes results in diminished cellular growth [@RN1755; @RN1754; @RN884].  Formation of new ribosomes is a complex and energy demanding process involving the whole transcriptional apparatus [@RN1920, @RN1820].  In skeletal muscle, a single bout of resistance exercise elevates markers of ribosome biogenesis in skeletal muscle evident from both Figueiredo et al. and Nader et al. who evaluated immature forms of rRNA in response to exercise in the untrained state [@RN1644; @RN1037]. Studies’ examining repeated bouts of resistance exercise suggests that RT leads to accumulation of ribosomes. Bickel et al. showed that two subsequent bouts of a hypertrophic stimulus increased total RNA at 72 h after a second bout of RT [@RN1520]. Following three weeks of RT, Brook et al. reported an increase in total RNA with no further increase after three additional weeks of training [@RN1642]. The findings of Brook et al. fits with observations from our laboratory where we saw a ~30% increase in total RNA per tissue weight after two weeks of RT. Following an additional 10 weeks of RT the RNA content tended to normalized without reaching baseline values (Hammarström et al. 2017 unpublished observation). These data suggests that ribosome accumulation per tissue weight, reaches a plateau in the early phase of RT. Interestingly, in the study by Figueiredo and colleagues, acute exercise in the trained state (after eight weeks of RT), rRNA and pre-rRNA rapidly decreases in response to acute exercise [@RN1644]. 
Based on these observations we are hypothesizing that (1) ribosome accumulation occurs in the early phase (3-4 weeks) of RT and (2) reaches a plateau during this phase in response to constant volume RT and finally (3) that fluctuations in ribosome abundance occurs in response to altered training volume and training cessation. Subsequently the aim of this study is to characterize fluctuations in ribosomal biogenesis during the initial phase of RT, in response to altered training volumes and in response to a short-term de-training period.

>>>>

Our skeletal muscles not only constitutes our locomotive apparatus but also serve as our primary storage site of amino-acids. This makes the skeletal muscle system a critical target for interventions that promotes health across the lifespan [@RN763]<!-- Wolfe 2006 -->.
The beneficial effects of resistance-training (RT) are in large part mediated through exercise-induced changes in the balance between muscular protein breakdown and synthesis, with successful RT leading to net protein accretion and muscle hypertrophy. A traditional view of exercise-induced muscle hypertrophy has been that muscle protein accumulates as a consequence of repeated episodes of elevated post-exercise protein synthesis in the acute phase, resulting in a net-positive protein balance over time. Indeed, one bout of resistance exercise acutely up-regulates skeletal muscle protein synthesis up to 48-hr after exercise [@RN786]<!-- Phillips 1997 -->
and subsequently, repeated bouts would lead to accumulation of muscle protein over time
[@RN1272<!-- Phillips review 2014 -->;
@RN1642]<!-- Brook 2015 experiment -->. 
In recent years, this view has been supplemented with evidence showing that chronic RT also leads to an increased resting synthetic rate of muscle protein
[@RN1521;<!-- Kim 2005 Fasted state skeletal muscle PS altered with training  --> @RN1866;<!-- Wilkingson 2008 -->
@RN1897]<!-- Reidy 2017 Increased rested state protein synthesis -->.
An increase in the resting protein synthetic rate has been postulated to be related to an increased translational capacity, i.e. accumulation of ribosomes
[@RN1897;<!-- Reidy 2017 -->
@RN1912]<!-- Figueiredo 2017 Ribosome biogenesis review -->,
a notion supported by observations relating exercise-induced changes in total RNA, a proxy marker for ribosome abundance, to muscle hypertrophy
[@RN1644;<!-- Figueiredo 2015 Correlation between total RNA fold change and CSA hyp -->@RN1755; <!-- Stec Ribosome bio higher in high responders -->
@Hammarstrm2020]<!-- Volume dependent RNA accum relates to volume depend hyper -->. Furthermore, inhibition of ribosomal RNA (rRNA) transcription or inhibition of up-stream transcription factors leads to halted biosynthesis of ribosomes resulting in diminished cellular growth
[@RN1755<!-- Stec 2016 Pol I-specific inhibition leads to GF induced myotube hypertrophty -->;
@RN1754<!-- West 2016 Inhibition of c-Myc (through inibitor) reduces RNA and MPS-->;
@vonWalden2016]<!-- Inhibition of Pol 1 rDNA tracsription and myotube growth -->.

Formation of new ribosomes is a complex and energy demanding process involving the whole transcriptional apparatus
[@RN1920<!-- Warner review yeast -->,
@RN1820;<!-- Moss review -->
@vonWalden2016]<!-- Inhibition of Pol II affects trasncription at Pol I -->.
In skeletal muscle, a single bout of resistance exercise elevates markers of ribosome biogenesis in skeletal muscle evident from studies investigating immature forms of rRNA in response to exercise in the untrained state
[@REFNEEDED<!-- References on acute 45S changes -->;
@RN1037<!-- Nader 2014 45S up in response to 4 h after acute exercise -->].
Studies examining repeated bouts of resistance exercise suggests that RT leads to accumulation of ribosomes. Bickel *et al*. showed that two subsequent bouts of a hypertrophic stimulus increased total RNA at 72 h after a second bout of RT
[@RN1520]<!-- Total RNA increase 72 h after a secound bout -->.
Following three weeks of RT, Brook et al. reported an increase in total RNA with no further increase after three additional weeks of training
[@RN1642]<!-- Brook 2015 6 week time-course -->.
The findings of Brook et al. fits with observations from our laboratory where we saw a ~30% increase in total RNA per tissue weight after two weeks of RT. Following an additional 10 weeks of RT the RNA content tended to normalized without reaching baseline values (Hammarström et al. 2017 unpublished observation). These data suggests that ribosome accumulation per tissue weight, reaches a plateau in the early phase of RT. Interestingly, in the study by Figueiredo and colleagues, acute exercise in the trained state (after eight weeks of RT), rRNA and pre-rRNA rapidly decreases in response to acute exercise [@RN1644]. 
Based on these observations we are hypothesizing that (1) ribosome accumulation occurs in the early phase (3-4 weeks) of RT and (2) reaches a plateau during this phase in response to constant volume RT and finally (3) that fluctuations in ribosome abundance occurs in response to altered training volume and training cessation. Subsequently the aim of this study is to characterize fluctuations in ribosomal biogenesis during the initial phase of RT, in response to altered training volumes and in response to a short-term de-training period.
